Precise Data Flow Analysis\\in the Presence of Correlated Method Calls\footnote{This research was supported by the Natural Sciences and Engineering Research Council of Canada and the Ontario Ministry of Research and Innovation.}

### 2015-06-19 sas2015_final/content/example.tex
creating final version without comments and redundant files

Figure~\ref{list:ccexample} shows a small Java program.
%that we will use to illustrate
%how our technique improves the precision of a simple taint analysis.
The program
declares a class \code{A} with a subclass \code{B},  where \code{A} defines methods 
\code{foo()} and \code{bar()} that are overridden in \code{B}.  
%To keep the 
%example small and self-contained, it is 
We assume  that secret values are created by 
an unspecified function \code{secret()}, which is called in \code{A.foo()} on 
line~\ref{line:Afoo}. 
%Furthermore, it is assumed that 
Any write to standard output 
%via calls \code{System.out.println()} such as the one in method \code{B.bar()} 
%on line~\ref{line:Bbar} 
is assumed to be a sink (e.g., the call to \code{System.out.println()} in \code{B.bar()}).
%a security-sensitive operation.
Depending on the number of 
arguments passed to the program, the \code{main()} method of the example program creates 
either an \code{A}-object or a \code{B}-object%
%on line~\ref{line:createObject}
. The 
program then calls \code{foo()} on this object
on line~\ref{line:callfoo}, which is 
followed by a call to \code{bar()} on the same object%
%on line~\ref{line:callbar}
.

%The specific question that we would like to answer for the example program of
%Figure~\ref{list:ccexample} is the following:
We wish to answer the following question: Is it possible for the untrusted value 
that is read on line~\ref{line:Afoo} to flow to the print statement% %on line~\ref{line:Bbar}
?
%To determine whether such a flow of tainted data is possible,
Consider the control-flow
supergraph for the example program that is shown in Figure~\ref{fig:examplesupergraph}.
 The nodes in this graph correspond to statements, method entry points (start nodes) and 
method exit points (end nodes). For each method call, the graph contains a 
distinct call-node and a return-node. Edges in the graph reflect intraprocedural control flow, 
flow of control from a caller to a callee (edges from call-nodes to start-nodes), or
flow of control from a callee back to a caller (edges from end-nodes to return-nodes).

@@

\commentout{
In summary, the example discussed in this section shows how the precision of 
IFDS-based taint analysis can be improved by taking into account how paths
become infeasible in the presence of correlated calls. In particular, we have
shown how a false positive reported by a standard IFDS-based algorithm is eliminated.
}
While the discussion in this section has focused on the specific problem of taint analysis,
our technique generally applies to \textit{any}
data-flow-analysis problem that can be expressed in the IFDS framework. This includes
many common analysis tasks such as reaching definitions, constant propagation, slicing,
typestate analysis, pointer analysis, and lightweight
shape analysis.

@@

In these programs, on average, 3\% of all call sites $C$ are polymorphic call sites $C_P$.
%\mtodo{Should we explain in a footnote why the number of polymorphic call sites is so low?}
%\otodo{We don't have an explanation.}
Out of these polymorphic call sites, a significant fraction (39\%) are correlated 
call sites $C^\Subset$. We also see that, on average,  each correlated-call receiver is involved in approximately
three correlated calls.

\subsection{An Example from the Scala Collections Library}\label{sec:collex}
The Scala collections library contains the trait \code{TraversableOnce} that is
shared by both collections and iterators over them. 
The \code{toArray} method of this trait
creates an array and copies the contents of the collection or iterator into it:
\inputMintedNoLNs{scala}{collectionsExample.scala}
%\otodo{Marianna, can you figure out how to get LaTeX to syntax-highlight this?} 
When \code{this} refers to an iterator rather than a collection, the call to \code{this.size}
extracts all elements of the iterator to count them.
 At the call to \code{copyToArray},
the iterator is already empty, so nothing is copied to the newly created array.
One could design an IFDS analysis to detect this kind of bug.

However, the implementation of \code{TraversableOnce.toArray} is actually correct because the above
code is guarded with a test: \code{if (this.isTraversableAgain) ...}
%\inputMintedNoLNs{scala}{isTraversable.scala}
When the \code{isTraversableAgain} method returns false, as it does for
an iterator, the \code{toArray} method uses a different
(less efficient) implementation. The bug report would therefore be a false positive. 
The \code{isTraversableAgain} method is easy to analyze: it returns the constant
true in a collection and the constant false in an iterator. However, in order
to eliminate the false positive bug report, an analysis would need to rule out infeasible
paths using correlated calls. Specifically, the following path triggers the bug, but
is infeasible: first, call \code{isTraversableAgain} on a collection, returning
true, then call \code{size} and \code{copyToArray} on an iterator. Our correlated
calls analysis could determine that this path is infeasible because it calls the
collection version of \code{isTraversableAgain} but the iterator versions of
\code{size} and \code{copyToArray}.
The relevant code from \textsf{TraversableOnce} and other related traits is shown
\techreport{in Figure~\ref{fig:traversableonce} }in the \reportOrAppendix.

>>>>

Figure~\ref{list:ccexample} shows a small Java program.
The program
declares a class \code{A} with a subclass \code{B},  where \code{A} defines methods 
\code{foo()} and \code{bar()} that are overridden in \code{B}.  
We assume  that secret values are created by 
an unspecified function \code{secret()}, which is called in \code{A.foo()} on 
line~\ref{line:Afoo}. 
Any write to standard output 
is assumed to be a sink (e.g., the call to \code{System.out.println()} in \code{B.bar()}).
Depending on the number of 
arguments passed to the program, the \code{main()} method of the example program creates 
either an \code{A}-object or a \code{B}-object. The 
program then calls \code{foo()} on this object
on line~\ref{line:callfoo}, which is 
followed by a call to \code{bar()} on the same object.

We wish to answer the following question: Is it possible for the untrusted value 
that is read on line~\ref{line:Afoo} to flow to the print statement?
Consider the control-flow
supergraph for the example program that is shown in Figure~\ref{fig:examplesupergraph}.
 The nodes in this graph correspond to statements, method entry points (start nodes) and 
method exit points (end nodes). For each method call, the graph contains a 
distinct call-node and a return-node. Edges in the graph reflect intraprocedural control flow, 
flow of control from a caller to a callee (edges from call-nodes to start-nodes), or
flow of control from a callee back to a caller (edges from end-nodes to return-nodes).

@@

While the discussion in this section has focused on the specific problem of taint analysis,
our technique generally applies to \textit{any}
data-flow-analysis problem that can be expressed in the IFDS framework. This includes
many common analysis tasks such as reaching definitions, constant propagation, slicing,
typestate analysis, pointer analysis, and lightweight
shape analysis.

@@

In these programs, on average, 3\% of all call sites $C$ are polymorphic call sites $C_P$.
Out of these polymorphic call sites, a significant fraction (39\%) are correlated 
call sites $C^\Subset$. We also see that, on average,  each correlated-call receiver is involved in approximately
three correlated calls.

\subsection{An Example from the Scala Collections Library}\label{sec:collex}
The Scala collections library contains the trait \code{TraversableOnce} that is
shared by both collections and iterators over them. 
The \code{toArray} method of this trait
creates an array and copies the contents of the collection or iterator into it:
\inputMintedNoLNs{scala}{collectionsExample.scala}
When \code{this} refers to an iterator rather than a collection, the call to \code{this.size}
extracts all elements of the iterator to count them.
 At the call to \code{copyToArray},
the iterator is already empty, so nothing is copied to the newly created array.
One could design an IFDS analysis to detect this kind of bug.

However, the implementation of \code{TraversableOnce.toArray} is actually correct because the above
code is guarded with a test: \code{if (this.isTraversableAgain) ...}
When the \code{isTraversableAgain} method returns false, as it does for
an iterator, the \code{toArray} method uses a different
(less efficient) implementation. The bug report would therefore be a false positive. 
The \code{isTraversableAgain} method is easy to analyze: it returns the constant
true in a collection and the constant false in an iterator. However, in order
to eliminate the false positive bug report, an analysis would need to rule out infeasible
paths using correlated calls. Specifically, the following path triggers the bug, but
is infeasible: first, call \code{isTraversableAgain} on a collection, returning
true, then call \code{size} and \code{copyToArray} on an iterator. Our correlated
calls analysis could determine that this path is infeasible because it calls the
collection version of \code{isTraversableAgain} but the iterator versions of
\code{size} and \code{copyToArray}.
The relevant code from \textsf{TraversableOnce} and other related traits is shown
\techreport{in Figure~\ref{fig:traversableonce} }in the \reportOrAppendix.
